---
title: "COMPSCIX 415.2 Homework 5"
author: "Jennifer Lu"
date: "7/13/2018"
output:
  html_document:
    toc: true
    theme: united
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(nycflights13)
```

### RStudio and R Markdown (3 points)
#### 1. Use markdown headers in your document to clearly separate each midterm question and add a table of contents to your document.

### The tidyverse packages (3 points)
By now you’ve used at least five different packages from the tidyverse for plotting, data munging, reshaping
data, importing/exporting data, and using tibbles (the tibble package is used for this without you even
realizing it’s there).

#### 1. Can you name which package is associated with each task below?
Plotting -
Data munging/wrangling -
Reshaping (speading and gathering) data -
Importing/exporting data -

#### 2. Now can you name two functions that you’ve used from each package that you listed above for these
tasks?
Plotting -
Data munging/wrangling -
Reshaping data -
Importing/exporting data -