---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Jennifer Lu"
date: "7/13/2018"
output:
  html_document:
    toc: true
    theme: united
    df_print: paged
---

My Github repository for my assignments can be found at this URL: [https://github.com/jencake/compscix-415-2-assignments](https://github.com/jencake/compscix-415-2-assignments)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(nycflights13)
```

## RStudio and R Markdown (3 points)
### 1. Use markdown headers in your document to clearly separate each midterm question and add a table of contents to your document.

## The tidyverse packages (3 points)
By now you’ve used at least five different packages from the tidyverse for plotting, data munging, reshaping
data, importing/exporting data, and using tibbles (the tibble package is used for this without you even
realizing it’s there).

### 1. Can you name which package is associated with each task below?
Plotting - *ggplot2*

Data munging/wrangling - *dplyr*

Reshaping (speading and gathering) data -*tidyr*

Importing/exporting data - *readr*

### 2. Now can you name two functions that you’ve used from each package that you listed above for these tasks?
Plotting - *ggplot() geom_bar()*

Data munging/wrangling - *filter(), summarize()*

Reshaping data - *gather(), spread()*

Importing/exporting data - *write_csv() read_csv()*

## R Basics (1.5 points)
### 1. Fix this code with the fewest number of changes possible so it works:
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
```

### 2. Fix this code so it works:
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

### 3. Look at the code below and comment on what happened to the values in the vector.
```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

The output type of c() is determined from the highest type of the componenets in the hierarchy NULL < raw < logical < integer < double < complex < character < list < expression. Therefore, the presence of characters along with integers transforms all elements of the output list to characters.

## Data import/export (3 points)
### 1. Download the rail_trail.txt file from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.

```{r}
rail_trail_fr_txt_file <- read_delim(file = 'rail_trail.txt', delim='|')
glimpse(rail_trail_fr_txt_file)
```

### 2. Export the file into a comma-separated file and name it “rail_trail.csv”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.
```{r}
write_csv(rail_trail_fr_txt_file, path = 'rail_trail.csv')
rail_trail_from_csv <- read_csv('rail_trail.csv')
glimpse(rail_trail_from_csv)
```

## Visualization (6 points)
### 1. Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.

Blue and pink would be clearer colors for the men vs women part of the graphic. The graphic would be more clear if it were separated into two plots, one with age group vs percentage of people and another with gender vs percentage of people. The title of the graph is "Percentage of respondents who say it is likely that a woman will be president in their lifetime," but the graph contains both "yes" and "no" proportion.

### 2. Reproduce this graphic using the diamonds data set.
```{r}
ggplot(data = diamonds, mapping = aes(x=cut, y=carat)) +
  geom_boxplot(mapping = aes(fill=color), position = 'identity') +
  coord_flip() +
  xlab('CUT OF DIAMOND') +
  ylab('CARAT OF DIAMOND')
```

### 3. The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.
```{r}
ggplot(data = diamonds, mapping = aes(x=cut, y=carat)) +
  geom_boxplot(mapping = aes(fill=color)) +
  coord_flip() +
  xlab('CUT OF DIAMOND') +
  ylab('CARAT OF DIAMOND')
```

## Data munging and wrangling (6 points)
### 1. Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.
```{r}
table2 %>%
  spread(key=type, value=count) %>%
  mutate(rate = cases / population * 10000)
```

### 2. Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.
```{r}
diamonds_with_price_per_carat <- diamonds %>%
  mutate(price_per_carat = price / carat)
```

### 3. For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.
```{r}
diamonds %>%
  group_by(cut) %>%
  summarize(
    count = n(),
    big_n_expensive = sum(price>10000 & carat<1.5),
    proportion = sum(price>10000 & carat<1.5)/n() 
  )
```

- Do the results make sense? Why?

The results make sense because you'd expect the demand for bigger and more expensive diamonds to increase as the cut gets better. Also, it's reasonable that the proportion of diamonds that cost $10k and is bigger than 1.5 carats is fairly small compared to the total number of diamonds manufactured.

- Do we need to be wary of any of these numbers? Why?

The proportions do not have equal predictive power for each cut because the number of diamonds manufactured by cut is different.

## EDA (6 points)
Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:

### 1. During what time period is this data from?
```{r}
txhousing %>%
  group_by(year) %>%
  summarize(
    min_month = min(month),
    max_month = max(month)
  )
```

The data spans between January 2000 and July 2015.

### 2. How many cities are represented?
```{r}
txhousing %>%
  summarize(
    cities = n_distinct(city)
  )
```

46 cities are represented.

### 3. Which city, month and year had the highest number of sales?
```{r}
txhousing %>%
  group_by(city, month, year) %>%
  summarize(
    most_sales = max(sales)
  ) %>%
  arrange(desc(most_sales))
```

The highest number of sale were in Houston, July 2015.

### 4. What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.

I think they're positively correlated.
```{r}
ggplot(data=txhousing, mapping = aes(x=listings, y=sales)) +
  geom_point()
```

### 5. What proportion of sales is missing for each city?
```{r}
txhousing %>%
  group_by(city) %>%
  summarize(
    proportion_missing = sum(is.na(sales))/n()
  )
```

### 6. Looking at only the cities and months with greater than 500 sales:
```{r}
more_than_500_sales <- txhousing %>%
  filter(sales>500)
```

- Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.
```{r}
more_than_500_sales %>%
  group_by(city) %>%
  summarize(
    min_median = min(median),
    max_median = max(median),
    spread = max(median) - min(median)
  ) %>%
  arrange(desc(max_median, min_median))

```

```{r}
ggplot(data=more_than_500_sales, mapping=aes(x=city, y=median)) +
  geom_boxplot() +
  coord_flip()
```

The median sale price distribution is different for each city.

- Any cities that stand out that you’d want to investigate further?

Corpus Christi seem to have very little median price variation year-round.

- Why might we want to filter out all cities and months with sales less than 500?

We only want to look at median prices in major cities during popular sale months.

## Git and Github (1.5 points)
To demonstrate your use of git and Github, at the top of your document put a hyperlink to your Github
repository.
Once you are finished with your midterm, commit your final changes with the comment “finished the midterm
- woohoo” and push your R Markdown file and your html or pdf file to Github.